Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-05-08 Manuscript/manuscript.Rmd
added class imbalance references in results

We developed a computational framework<!--, refered to as *TADRF*,--> aimed at determining the optimal set of data level characteristics necessary for topologically associated domain (TAD) boundary prediction. We chose the random forest (RF) machine learning algorithm as our binary classification tool. The reason for choosing RF was two-fold: (1) In order to devise a tunable prediction rule in a supervised learning framework that was both robust to overfitting and able to handle multiple correlated predictors, and (2) to allow for an interpretable ranking of predictors with respect to their ability to predict boundary regions [@boulesteix2012overview] [??? Read 10.1002/widm.1072 and 10.1002/widm.1301, elaborate, cite;!!!added first source]. We then leveraged our optimally built RF to precisely predict the genomic coordinates of both TAD and loop boundaries at base pair resolution.

@@

Boundary regions were defined as regions containing a called boundary ($Y = 1$), while non-boundary regions were defined as genomic bins that did not contain a boundary ($Y = 0$) (Figure 1A, see Methods). The total number of TADs, TAD boundaries, and the number of genomic bins expectedly decreased with the decreased resolution of Hi-C data (Table 1, Supplementary Table 3). The number of non-boundary regions highly outnumbered boundary regions, and these results were consistent for loop boundaries (Supplementary Table 4). Such disproportional presence of examples in one class is known as "class imbalance" problem that negatively affects model performance [??? Add refs]. To address class imbalance, we evaluated the effect of three re-sampling techniques on prediction performance. *Random over-sampling* (ROS) was defined as sampling with replacement from boundary regions. *Random under-sampling* (RUS) was defined as sampling with replacement from non-boundary regions. Lastly, we considered *Synthetic minority over-sampling technique* (SMOTE), which is a combination of both random over- and under-sampling [@chawla2002smote] (see Methods). These strategies allowed us to select the best performing model.

>>>>

We developed a computational framework<!--, refered to as *TADRF*,--> aimed at determining the optimal set of data level characteristics necessary for topologically associated domain (TAD) boundary prediction. We chose the random forest (RF) machine learning algorithm as our binary classification tool. The reason for choosing RF was two-fold: (1) In order to devise a tunable prediction rule in a supervised learning framework that is both robust to overfitting and able to handle multiple correlated predictors, and (2) to allow for an interpretable ranking of predictors with respect to their ability to predict boundary regions [@boulesteix2012overview] [??? Read 10.1002/widm.1072 and 10.1002/widm.1301, elaborate, cite;!!!added first source]. We then leveraged our optimally built RF to precisely predict the genomic coordinates of both TAD and loop boundaries at base pair resolution.

@@

Boundary regions were defined as regions containing a called boundary ($Y = 1$), while non-boundary regions were defined as genomic bins that did not contain a boundary ($Y = 0$) (Figure 1A, see Methods). The total number of TADs, TAD boundaries, and the number of genomic bins expectedly decreased with the decreased resolution of Hi-C data (Table 1, Supplementary Table 3). The number of non-boundary regions highly outnumbered boundary regions, and these results were consistent for loop boundaries (Supplementary Table 4). Such disproportional presence of examples in one class is known as a "class imbalance" problem that negatively affects model performance [@prati2011survey; @wei2013role] [??? Add refs; !!!added]. To address class imbalance, we evaluated the effect of three re-sampling techniques on prediction performance. *Random over-sampling* (ROS) was defined as sampling with replacement from boundary regions. *Random under-sampling* (RUS) was defined as sampling with replacement from non-boundary regions. Lastly, we considered *Synthetic minority over-sampling technique* (SMOTE), which is a combination of both random over- and under-sampling [@chawla2002smote] (see Methods). These strategies allowed us to select the best performing model.
